---
title: "Mental Health and Standardized Test Scores"
---

## Motivation and Context

```{r}
#| label: do this first
#| echo: false
#| message: false

# change this to the location of your Quarto file containing your project; then delete this comment
here::i_am("math-437-project-template.qmd")
```

It is commonly said that America today is in a mental health crisis - the CDC states that a staggering one in five Americans suffer from anxiety or depression, with the youth being even more severely effected [https://www.cdc.gov/mental-health/about/what-cdc-is-doing.html]. With the increase in the severity in this crisis has apparently come an increase in awareness and resources available to help meet it, however. In this project I intend to only cover mental health issues and educational results that are common and familiar to the average person (i.e., depression and anxiety for mental health, and standardized test scores for educational results). Thus, no additional context is required for a general audience to understand the concepts at play. 

## Main Objective

## Packages Used In This Analysis

```{r}
#| label: load packages
#| message: false
#| warning: false

library(here)
library(readr)
library(dplyr)
library(ggplot2)
library(rsample)

library(readxl)
```


| Package | Use |
|-------------------------------|----------------------------------------|
| [here](https://github.com/jennybc/here_here) | to easily load and save data |
| [readr](https://readr.tidyverse.org/) | to import the CSV file data |
| [dplyr](https://dplyr.tidyverse.org/) | to massage and summarize data |
| [rsample](https://rsample.tidymodels.org/) | to split data into training and test sets |
| [ggplot2](https://ggplot2.tidyverse.org/) | to create nice-looking and informative graphs |

## Data Description

```{r}
#| label: import data
#| warning: false
test_scores <- read_excel(here::here("Data/standardized_test_scores.xlsx"))
sad_percent <- read_excel(here::here("Data/felt_sad_2021.xlsx"))
```


```{r}
#| label: putting the data together
data <- test_scores[4:54, 1:3]
data[, 4] <- sad_percent[5:55, 2]
colnames(data)[4] <- "Sadness"

data
```

### Data Limitations

## Data Wrangling (Optional Section)

## Exploratory Data Analysis

## Modeling

## Insights

### Limitations and Future Work

### Reflection (Optional Subsection)

(Reflection included in Canvas submission)
