---
title: "Mental Health and Standardized Test Scores"
---

## Motivation and Context

```{r}
#| label: do this first
#| echo: false
#| message: false

# change this to the location of your Quarto file containing your project; then delete this comment
here::i_am("math-437-project-template.qmd")
```

## Main Objective

## Packages Used In This Analysis

```{r}
#| label: load packages
#| message: false
#| warning: false

library(here)
library(readr)
library(dplyr)
library(ggplot2)
library(rsample)
```


| Package | Use |
|-------------------------------|----------------------------------------|
| [here](https://github.com/jennybc/here_here) | to easily load and save data |
| [readr](https://readr.tidyverse.org/) | to import the CSV file data |
| [dplyr](https://dplyr.tidyverse.org/) | to massage and summarize data |
| [rsample](https://rsample.tidymodels.org/) | to split data into training and test sets |
| [ggplot2](https://ggplot2.tidyverse.org/) | to create nice-looking and informative graphs |

## Data Description

```{r}
#| label: import data
#| warning: false
test_scores <- read_excel(here::here("Data/standardized_test_scores.xlsx"))
sad_percent <- read_excel(here::here("Data/felt_sad_2021.xlsx"))
```


```{r}
#| label: putting the data together
data <- test_scores[4:54, 1:3]
data[, 4] <- sad_percent[5:55, 2]
colnames(data)[4] <- "Sadness"

data
```

### Data Limitations

## Data Wrangling (Optional Section)

## Exploratory Data Analysis

## Modeling

## Insights

### Limitations and Future Work

### Reflection (Optional Subsection)
